Enhancing \Hinf{} Norm Estimation using Local LPM/LRM Modeling: Applied to an AVIS\thanksref{thx}

### 2016-06-24 ch03-lrm/nonparametric.tex
Some recommendations private defense
Carry out some small changes

\section{Introduction}
\label{sec:nonparametric:introduction}

@@

\subsection{The \glsentrydesc{LRM}}
\label{sec:nparam:LRM}
The \gls{LRM} as first introduced by \citet{McKelvey2012LRM}, overcomes the computational burden of an iterative procedure by weighting the equation error by the denominator polynomial $\LocalModel{A}$  akin to the procedure in \citep{Levy1959}.
I.e. the \gls{LRM} procedure tries to minimize the equation error in
\begin{equation}
  \LocalModel{A} \hadamard \LocalVector{Y} = \LocalModel{B} \hadamard \LocalVector{U}  + \LocalVector{C} + \LocalVector{V}
  \text{,}
\end{equation}
for which the corresponding cost function is
\begin{equation}
  \LocalVector{\CostFunc{\LRM}}
  \isdef 
  \left( \LocalModel{A} \hadamard \LocalVector{Y}  -  \LocalModel{B} \hadamard \LocalVector{U} - \LocalModel{C} \right)^{\HT} 
      \left( \LocalModel{A} \hadamard \LocalVector{Y}  -  \LocalModel{B} \hadamard \LocalVector{U} - \LocalModel{C} \right)
      \text{.}
      \label{eq:nparam:LRM:costFunc}
\end{equation}
where $\LocalVector{V}$ is vector consisting of \gls{iid} complex normally distributed variables~\citep{Gallager2008} that each have a variance $\sigma_V^2$, i.e. the disturbing noise is assumed white over the local frequency window.
Equivalently, the last equation can be rewritten as a linear regression problem
\begin{equation}
  \LocalVector{Y} = \LocalMatrix{K} \LocalVector{\theta} + \LocalVector{V}
  \label{eq:nparam:linearRegression}
\end{equation}
where $\LocalMatrix{K}$ is the so-called design matrix (or observation matrix):
\begin{align}
  \label{eq:nparam:designMatrix}
  \LocalMatrix{K} 
    & \isdef 
    \phantom{+}
  \mat{
     \LocalMatrix{K}_A &
     \LocalMatrix{K}_B & 
     \LocalMatrix{K}_C
  }\\
  \LocalMatrix{K}_A 
    &\isdef
    \phantom{+}
    \mat{
      \LocalVector{Y} \hadamard \LocalVector{d}^1 &
      \cdots &
      \LocalVector{Y} \hadamard \LocalVector{d}^{\order{A}}
    }\\
  \LocalMatrix{K}_B 
    &\isdef
    - \mat{
      \LocalVector{U} \hadamard \LocalVector{d}^0 &
      \cdots &
      \LocalVector{U} \hadamard \LocalVector{d}^{\order{B}}
    }\\
  \LocalMatrix{K}_C
    &\isdef
    - \mat{
      \LocalVector{d}^0 &
      \cdots &
      \LocalVector{d}^{\order{C}}
    }
    \text{.}
\end{align}
In this formulation, we have used $\LocalVector{d}^{n}$ to denote the $n^{\text{th}}$ Hadamard power of $\LocalVector{d}$ (this corresponds to \mcode{d.^n} in \MATLAB) such that
\begin{equation}
    \LocalVector{d}^{n} 
    \isdef
    \mat{
      (-\numel{W})^{n} &
      \cdots &
      (\numel{W})^{n}
    }^{\T}
    \text{ with $n \in \mathbb{Z}$}
\end{equation}
where every element corresponds to a value of $d^{n}$ in accordance with $\LocalShifts{k}$.

@@

For linear models (i.e. $\LocalVector{Y} = \LocalMatrix{K} \LocalVector{\theta}$, such as for the \gls{LRM} and \gls{LPM}), however, the \gls{PRESS} can be calculated in a single step without estimating any additional models~\citep[Sec.~12.3.2]{Seber2003}:
\begin{equation}
\PRESS = \frac{\LocalVector{E}^{\HT} \LocalVector{W}^{-1} \LocalVector{E}}{2N_W + 1}
\end{equation}
where $\LocalVector{E}$ is the residual vector of the full estimation and $\LocalVector{W}$ is a diagonal weighting matrix with $\LocalVector{W}_{ii} = (1 - \LocalVector{H}_{ii})^{\HT} (1 - \LocalVector{H}_{ii})$ and
$\LocalVector{H} = \LocalMatrix{K} \pinv{\LocalMatrix{K}}$ is the so-called `hat-matrix' that is known to be idempotent.

@@

When an optimal model order is to be selected for the whole frequency range, the local optimality criteria need to be reformulated in the global context of all local models over the frequency.
In practice, the formulas of the previous section extend quite ``naturally'' to this global context.
Particularly for the \gls{PRESS}:
\begin{equation}
  \PRESS_{\mathrm{global}} \propto \sum_k \PRESS_k
\end{equation}
where $\PRESS_k$ denotes the \gls{PRESS} for the $k^{\text{th}}$ local model.
One should realize that for all local models together ($M =  \cup_k \LocalModel[k]{M}$) and the global dataset $Z$, the global log-likelihood, cost function and number of parameters can be related to their local counterparts:
\begin{align}
  \loglikelihood{M | Z}   &= \sum_k \loglikelihood{\LocalModel[k]{M} | \LocalVector{Z}_{k}} \\
  L_{\mathrm{global}} &= \sum_k \LocalVector{L}_k\\
  \nth^{\mathrm{global}} &= \sum_k \nth(k)
  \text{.}
\end{align}
These global quantities can be substituted in the expressions of the local \gls{AIC} and \gls{AICc} of the previous section. 
Note that one should also $\nWind$ as the total number of data points (and not the number of points in each window).

@@

For the low \gls{SNR} case (top plots in  \figref{fig:nparam:PRESS} ), it can be seen that both \gls{LPM} and \gls{LRM} result in a very similar estimate of the noise level.
The values of the \gls{PRESS} are again very noisy, especially for the \gls{LPM} estimates.
Near the resonance, the $\lrm{7,1,1,1}$ results in the smalles \gls{PRESS} value.
This is in correspondence with the observations of the variance for similar noise levels in \figref{fig:nparam:comparison:terribleSNR}.
Note however that the results do not carry over to the \gls{RMS} error: since the \gls{LRM} is severely biased, the \gls{PRESS} becomes unreliable to detect model errors.

>>>>

\section{Introduction}
\label{sec:nonparametric:introduction}
%TODO: go back further in time for refs

@@

\subsection{The \glsentrydesc{LRM}}
\label{sec:nparam:LRM}
The \gls{LRM} as first introduced by \citet{McKelvey2012LRM}, overcomes the computational burden of an iterative procedure by weighting the equation error by the denominator polynomial $\LocalModel{A}$  akin to the procedure in \citep{Levy1959}.
I.e. the \gls{LRM} procedure tries to minimize the equation error in
\begin{equation}
  \LocalModel{A} \hadamard \LocalVector{Y} = \LocalModel{B} \hadamard \LocalVector{U}  + \LocalModel{C} + \LocalVector{V}
  \text{,}
\end{equation}
for which the corresponding cost function is
\begin{equation}
  \LocalVector{\CostFunc{\LRM}}
  \isdef 
  \left( \LocalModel{A} \hadamard \LocalVector{Y}  -  \LocalModel{B} \hadamard \LocalVector{U} - \LocalModel{C} \right)^{\HT} 
      \left( \LocalModel{A} \hadamard \LocalVector{Y}  -  \LocalModel{B} \hadamard \LocalVector{U} - \LocalModel{C} \right)
      \text{.}
      \label{eq:nparam:LRM:costFunc}
\end{equation}
where $\LocalVector{V}$ is vector consisting of \gls{iid} complex normally distributed variables~\citep{Gallager2008} that each have a variance $\sigma_V^2$, i.e. the disturbing noise is assumed white over the local frequency window.
Equivalently, the last equation can be rewritten as a linear regression problem
\begin{equation}
  \LocalVector{Y} = \LocalMatrix{K} \LocalVector{\theta} + \LocalVector{V}
  \label{eq:nparam:linearRegression}
\end{equation}
where $\LocalMatrix{K}$ is the so-called design matrix (or observation matrix):
\begin{align}
  \label{eq:nparam:designMatrix}
  \LocalMatrix{K} 
    & \isdef 
    \phantom{+}
  \mat{
     \LocalMatrix{K}_A &
     \LocalMatrix{K}_B & 
     \LocalMatrix{K}_C
  }\\
  \LocalMatrix{K}_A 
    &\isdef
    \phantom{+}
    \mat{
      \LocalVector{Y} \hadamard \LocalVector{d}^1 &
      \cdots &
      \LocalVector{Y} \hadamard \LocalVector{d}^{\order{A}}
    }\\
  \LocalMatrix{K}_B 
    &\isdef
    - \mat{
      \LocalVector{U} \hadamard \LocalVector{d}^0 &
      \cdots &
      \LocalVector{U} \hadamard \LocalVector{d}^{\order{B}}
    }\\
  \LocalMatrix{K}_C
    &\isdef
    - \mat{
      \LocalVector{d}^0 &
      \cdots &
      \LocalVector{d}^{\order{C}}
    }
    \text{.}
\end{align}
In this formulation, we have used $\LocalVector{d}^{n}$ to denote the $n^{\text{th}}$ Hadamard power of $\LocalVector{d}$ (this corresponds to \mcode{d.^n} in \MATLAB) such that
\begin{equation}
    \LocalVector{d}^{n} 
    \isdef
    \mat{
      (-\numel{W})^{n} &
      \cdots &
      (\numel{W})^{n}
    }^{\T}
    \text{ with $n \in \mathbb{Z}$}
\end{equation}
where every element corresponds to a value of $d^{n}$ in accordance with $\LocalShifts{k}$.

@@

For linear models (i.e. $\LocalVector{Y} = \LocalMatrix{K} \LocalVector{\theta}$, such as for the \gls{LRM} and \gls{LPM}), however, the \gls{PRESS} can be calculated in a single step without estimating any additional models~\citep{Golub1979}:
\begin{equation}
\PRESS = \frac{\LocalVector{E}^{\HT} \LocalVector{W}^{-1} \LocalVector{E}}{2N_W + 1}
\end{equation}
where $\LocalVector{E}$ is the residual vector of the full estimation and $\LocalVector{W}$ is a diagonal weighting matrix with $\LocalVector{W}_{ii} = (1 - \LocalVector{H}_{ii})^{\HT} (1 - \LocalVector{H}_{ii})$ and
$\LocalVector{H} = \LocalMatrix{K} \pinv{\LocalMatrix{K}}$ is the so-called `hat-matrix' that is known to be idempotent.
The proof for this can also be found in \citet[Sec.~12.3.2]{Seber2003}.

@@

When an optimal model order is to be selected for the whole frequency range, the local optimality criteria need to be reformulated in the global context of all local models over the frequency.
In practice, the formulas of the previous section extend quite ``naturally'' to this global context.
Particularly for the \gls{PRESS}:
\begin{equation}
  \PRESS_{\mathrm{global}} \propto \sum_k \PRESS_k
\end{equation}
where $\PRESS_k$ denotes the \gls{PRESS} for the $k^{\text{th}}$ local model.
One should realize that for all local models together ($M =  \cup_k \LocalModel[k]{M}$) and the global dataset $Z$, the global log-likelihood, cost function and number of parameters can be related to their local counterparts:
\begin{align}
  \loglikelihood{M | Z}   &= \sum_k \loglikelihood{\LocalModel[k]{M} | \LocalVector{Z}_{k}} \\
  L_{\mathrm{global}} &= \sum_k \LocalVector{L}_k\\
  \nth^{\mathrm{global}} &= \sum_k \nth(k)
  \text{.}
\end{align}
These global quantities can be substituted in the expressions of the local \gls{AIC} and \gls{AICc} of the previous section. 
Note that one should interpret $\nWind$ as the total number of data points (and not the number of points in each window) for the global case.

@@

For the low \gls{SNR} case (top plots in  \figref{fig:nparam:PRESS} ), it can be seen that both \gls{LPM} and \gls{LRM} result in a very similar estimate of the noise level.
The values of the \gls{PRESS} are again very noisy, especially for the \gls{LPM} estimates.
Near the resonance, the $\lrm{7,1,1,1}$ results in the smallest \gls{PRESS} value.
This is in correspondence with the observations of the variance for similar noise levels in \figref{fig:nparam:comparison:terribleSNR}.
Note however that the results do not carry over to the \gls{RMS} error: since the \gls{LRM} is severely biased, the \gls{PRESS} becomes unreliable to detect model errors.

### 2016-06-24 ch04-lrmhinf/lrmhinf.tex
Add some TODOs from Jury/self
Carry out some small changes

\section{Simulation}
\label{sec:lrmhinf:simulation}
To allow the generalization of the results of the simulations, we first introduce a way to decompose complex systems into simpler systems.
The behavior of these simpler systems then allows to predict the behavior for these more complex systems.
Concretely, any proper real-rational system $\Delta$ can be decomposed in a parallel connection of lower-order sub-systems by means of a partial fraction expansion as in \eqref{eq:excitation:PFE}, i.e.
\begin{equation}
\Delta(s) = 
         \deemph{
         \sum_{i=1}^{I_1} 
          \sum_{\multiplicity=1}^{M_{1,i}} 
            \frac{\dc[i]}
                      { \left( s - \pole[i] \right)^{\multiplicity}}
            + 
        }
        \sum_{i=1}^{I_2} 
          \deemph{\sum_{\multiplicity=1}^{M_{2,i}}}
           \frac{\dc[i,\multiplicity] \wn[i]^{2\multiplicity} \left( s + \wz[i] \right)}
                     {\wz[i] \left(s^2 + 2 \damping[i] \wn[i] s + \wn[i]^2 \right)^{\multiplicity}}
\end{equation}
with the same notations as in \chapref{sec:excitation}, most importantly, each resonant pole can be characterized by its frequency $\wn[i]$ and damping $\damping[i]$.
In the structural engineering community\citep[Section 2.2.2][see e.g.{Gawronski2004} it is well-known that around sharp resonance peaks ($\omega \approx \wn[i]$), the shape of a structural transfer function is mainly determined by its modal transfer function.
In many practical situations, this means that the gray terms in the equation often have a negligible influence.
For local modeling, the response is hence well-approximated by only considering the poles and zeros in the local frequency window.
To estimate $\infnorm{\Delta}$ mainly the resonant sub-systems are of interest as those are responsible for large gains of the system.
Assuming such resonances are sufficiently separated in the frequency,~\citep{Schoukens2013LPMerror,Geerardyn2013TIM,Gawronski2004} show that results for a single resonance are applicable to more complex systems.
This is done by considering only the dominant resonance -- \emph{viz.} the one that is the most relevant for the $\Hinf{}$ controller design.
By normalizing the conclusions for the time constant $\timeconst$ (or the $3\unit{dB}$ bandwidth), the results are generally valid for different values of the damping $\damping$.
Specifically, the time constant $\timeconst \isdef (\damping \wn)^{-1}$ of the dominant pole dictates the required measurement time~\citep{Schoukens2013LPMerror}, where $\damping$ is the relative damping and $\wn$ the natural frequency of the dominant pole.

>>>>

\section{Simulation}
\label{sec:lrmhinf:simulation}
To allow the generalization of the results of the simulations, we first introduce a way to decompose complex systems into simpler systems.
The behavior of these simpler systems then allows to predict the behavior for these more complex systems.
Concretely, any proper real-rational system $\Delta$ can be decomposed in a parallel connection of lower-order sub-systems by means of a partial fraction expansion as in \eqref{eq:excitation:PFE}, i.e.
\begin{equation}
\Delta(s) = 
         \deemph{
         \sum_{i=1}^{I_1} 
          \sum_{\multiplicity=1}^{M_{1,i}} 
            \frac{\dc[i]}
                      { \left( s - \pole[i] \right)^{\multiplicity}}
            + 
        }
        \sum_{i=1}^{I_2} 
          \deemph{\sum_{\multiplicity=1}^{M_{2,i}}}
           \frac{\dc[i,\multiplicity] \wn[i]^{2\multiplicity} \left( s + \wz[i] \right)}
                     {\wz[i] \left(s^2 + 2 \damping[i] \wn[i] s + \wn[i]^2 \right)^{\multiplicity}}
\end{equation}
with the same notations as in \chapref{sec:excitation}, most importantly, each resonant pole can be characterized by its frequency $\wn[i]$ and damping $\damping[i]$.
In the structural engineering community~\citep[see e.g.][Section 2.2.2]{Gawronski2004} it is well-known that around sharp resonance peaks ($\omega \approx \wn[i]$), the shape of a structural transfer function is mainly determined by its modal transfer function.
In many practical situations, this means that the gray terms in the equation often have a negligible influence.
For local modeling, the response is hence well-approximated by only considering the poles and zeros in the local frequency window.
To estimate $\infnorm{\Delta}$ mainly the resonant sub-systems are of interest as those are responsible for large gains of the system.
Assuming such resonances are sufficiently separated in the frequency,~\citep{Schoukens2013LPMerror,Geerardyn2013TIM,Gawronski2004} show that results for a single resonance are applicable to more complex systems.
This is done by considering only the dominant resonance -- \emph{viz.} the one that is the most relevant for the $\Hinf{}$ controller design.
By normalizing the conclusions for the time constant $\timeconst$ (or the $3\unit{dB}$ bandwidth), the results are generally valid for different values of the damping $\damping$.
Specifically, the time constant $\timeconst \isdef (\damping \wn)^{-1}$ of the dominant pole dictates the required measurement time~\citep{Schoukens2013LPMerror}, where $\damping$ is the relative damping and $\wn$ the natural frequency of the dominant pole.
